---
title: "Logistic Regression on Effective Treatments for Diabetes  "
output: html_notebook
---

Problem statement: Predicting effective treatments for diabetes in turn reducing the readmission into the hospital

Diabetes 130-US hospitals for years 1999-2008 Data Set

Source:

The data are submitted on behalf of the Center for Clinical and Translational Research, Virginia Commonwealth University, a recipient of NIH CTSA grant UL1 TR00058 and a recipient of the CERNER data. John Clore (jclore '@' vcu.edu), Krzysztof J. Cios (kcios '@' vcu.edu), Jon DeShazo (jpdeshazo '@' vcu.edu), and Beata Strack (strackb '@' vcu.edu). This data is a de-identified abstract of the Health Facts database (Cerner Corporation, Kansas City, MO).

Data Set Information:

The dataset represents 10 years (1999-2008) of clinical care at 130 US hospitals and integrated delivery networks. It includes over 50 features representing patient and hospital outcomes. Information was extracted from the database for encounters that satisfied the following criteria.
(1) It is an inpatient encounter (a hospital admission).
(2) It is a diabetic encounter, that is, one during which any kind of diabetes was entered to the system as a diagnosis.
(3) The length of stay was at least 1 day and at most 14 days.
(4) Laboratory tests were performed during the encounter.
(5) Medications were administered during the encounter.
The data contains such attributes as patient number, race, gender, age, admission type, time in hospital, medical specialty of admitting physician, number of lab test performed, HbA1c test result, diagnosis, number of medication, diabetic medications, number of outpatient, inpatient, and emergency visits in the year before the hospitalization, etc.
```{r}
d_data = read.csv('/Users/berkayozdin/Desktop/folders/dataset_diabetes/diabetic_data.csv', stringsAsFactors = FALSE)

str(d_data)
head(d_data)
```
Filtering patients with diabetes & patients who did not re-admit
```{r}
library(tidyverse)
d_data = d_data %>% filter(diabetesMed == 'Yes', readmitted == 'NO')

```
Excluding patients who are dead and are in hospise
```{r}
d_data = d_data %>% filter(discharge_disposition_id != '11' & discharge_disposition_id != '13' & discharge_disposition_id != '14' & discharge_disposition_id != '19' & discharge_disposition_id != '20')
```
Handling missing & erroneous values
```{r}
library(dplyr)
d_data = select(d_data, -c(weight,payer_code,medical_specialty))

d_data[d_data == "?"] = "Caucasian"
```
Custom encoding for the 23 drug features
```{r}

variables = c('metformin', 'repaglinide', 'nateglinide','chlorpropamide', 'glimepiride', 'acetohexamide', 'glipizide','glyburide', 'tolbutamide', 'pioglitazone', 'rosiglitazone', 'acarbose','miglitol', 'troglitazone', 'tolazamide', 'examide', 'citoglipton','insulin', 'glyburide.metformin', 'glipizide.metformin', 'glimepiride.pioglitazone', 'metformin.rosiglitazone', 'metformin.pioglitazone')
treatments = d_data[variables]
rownames(treatments) = d_data$encounter_id

treatments[treatments == "No"] = "0"
treatments[treatments == "Down"] = "1"
treatments[treatments == "Steady"] = "1"
treatments[treatments == "Up"] = "1"

treatments = treatments %>% mutate_at(variables, as.numeric)

```
Patients are Given at max a combination of 6 drugs for treating diabetes
```{r}

count_drugs = rowSums(treatments)
max = max(count_drugs)
paste('Max. # of drugs for treatment: ',max)
```
When the value of Insulin is '1' , creating the classes "insulin" & "insulin++" (insulin + others )**
```{r}
treatments$count_drugs = count_drugs
treatments$treatment = ifelse(treatments$insulin == 0 & count_drugs > 0 ,'other', 'insulin')

treatments$treatment <- ifelse(treatments$insulin==0, "other", ifelse(treatments$insulin==1 & count_drugs > 1, "insulin++", ifelse(treatments$insulin==1 & count_drugs ==1, "insulin",NA)))

```
Concat to actual dataframe
```{r}
d_data$treatments = treatments$treatment
```
Remove the drug and diagnosis columns
```{r}
d_data = d_data[ , !(names(d_data) %in% variables)]
diagnosis = c('diag_1','diag_2','diag_3')
d_data = d_data[ , !(names(d_data) %in% diagnosis)]
```
Choosing the records with insulin treatment
```{r}
d_data = d_data %>% filter(d_data$treatments != 'other')
```
Encoding the nominal categorical values & age column
```{r}
library(fastDummies)
dummies = c('race', 'gender','max_glu_serum', 'A1Cresult', 'change','diabetesMed', 'readmitted')
d_data <- dummy_cols(d_data, select_columns = dummies)
d_data = d_data[ , !(names(d_data) %in% dummies)]

to_change = c('1','2','3','4','5','6','7','8','9','10')
symbols = c("\\[0-10)","\\[10-20)","\\[20-30)","\\[30-40)","\\[40-50)","\\[50-60)","\\[60-70)","\\[70-80)","\\[80-90)","\\[90-100)")
for (i in 1:10) {
  d_data$age = gsub(symbols[i],to_change[i],d_data$age)
}
d_data$age = as.integer(d_data$age)
str(d_data)
```
Independence Tests
```{r}

test_columns = c('encounter_id', 'patient_nbr', 'age', 'admission_type_id',
       'discharge_disposition_id', 'admission_source_id', 'time_in_hospital',
       'num_lab_procedures', 'num_procedures', 'num_medications',
       'number_outpatient', 'number_emergency', 'number_inpatient','number_diagnoses',
       'race_AfricanAmerican', 'race_Asian', 'race_Caucasian', 'race_Hispanic',
       'race_Other', 'gender_Female', 'gender_Male',
       'max_glu_serum_>200', 'max_glu_serum_>300', 'max_glu_serum_None',
       'max_glu_serum_Norm', 'A1Cresult_>7', 'A1Cresult_>8', 'A1Cresult_None',
       'A1Cresult_Norm', 'change_Ch', 'change_No', 'diabetesMed_Yes',
       'readmitted_NO')

for (i in test_columns) {
  a = chisq.test(d_data[i],d_data$treatments)
  print(i)
  print(a)
 }

```
Logistic Regression
```{r}
insignificant_variables = c('encounter_id','patient_nbr','num_lab_procedures','number_outpatient','number_emergency','race_Asian','race_Other','diabetesMed_Yes','max_glu_serum_>200','A1Cresult_>8','A1Cresult_Norm')

d_data = d_data[ , !(names(d_data) %in% insignificant_variables)]

readmitted = d_data$readmitted_NO

d_data = d_data %>% mutate_if(is.numeric, funs(as.numeric(scale(.))))

d_data$readmitted_NO = readmitted

d_data$treatments = d_data$treatments %>% as.factor()

set.seed(1234)
create_train_test = function(data, size = 0.8, train = TRUE) {
    n_row = nrow(data)
    total_row = size * n_row
    train_sample <- 1: total_row
    if (train == TRUE) {
        return (data[train_sample, ])
    } else {
        return (data[-train_sample, ])
    }
}
data_train = create_train_test(d_data, 0.8, train = TRUE)
data_test = create_train_test(d_data, 0.8, train = FALSE)

formula = treatments~.
logit = glm(formula, data = data_train, family = 'binomial')
summary(logit)
```

